Palomar 5 Gaps

### 2024-10-04 paper/main.tex
retouch the overview again

    We applied the gap detection methods described in Appendix~\ref{sec:gap_detection}, along with the perturber identification methods from Appendix~\ref{sec:Perturber_Identification}. Results from all simulations are presented in Appendix~\ref{sec:gallery_of_gaps}. Once we identified the perturbers responsible for each gap, we conducted additional simulations to verify correct identification of the perturbers. These additional contain only the suspected clusters and were launched one-at-a-time to isolate the effects of a specific clusters. These results are shown in Fig.~\ref{fig:decomposition}.

    We present a summary of the number of gaps found in the simulation of the gallery of gaps from Appendix.~\ref{sec:gallery_of_gaps} in Fig.~\ref{fig:histogram_impact_time}, where the number of impacts and the time of impact is reported.

@@

    One striking result is that more than half of the gaps observed are caused by NGC2808. Upon examining the figures in the appendix, we note that these gaps are of similar size and occur in almost the same position in the leading tail across most simulations. This is no coincidence, as the size of a gap is a function of time, and recent encounters allow only limited time for the gap to develop. As shown in Fig.\ref{fig:histogram_impact_time}, the gaps caused by NGC2808 result from a flyby that occurred approximately 200 Myr ago. It is expected that a recent flyby appears in most simulations, as uncertainties grow when we look further back in time, allowing broader variation in the phase space positions of the clusters. However, due to relatively low uncertainties associated with NGC2808 and Palomar 5 compared to the integration time, the recent flyby of NGC~2808 occurs in 44 out of 50 simulations. Although the flybys do not occur identically in each simulation, the range in gap contrast, shown in the appendix, indicates that closer encounters lead to more significant impacts.

    Additionally, we observe that most gaps occur in the leading tail. We believe this to be a coincidence, as stars escape from the globular cluster's Lagrange points at roughly equal rates. The two main differences are the orbital phase and that the leading tail is about 400~pc closer to the Galactic center than the trailing tail. However, this difference is negligible compared to the orbital radius of 8-12kpc. As such, we expect the likelihood of a gap appearing in either tail to be equal, with the presence of one gap not influencing the probability of another.

    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation. However, 44 of these gaps are associated with NGC~2808. As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter. We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events. After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations. By integrating the binomial distribution, we compute that the probability of having at least 8 successes or fewer given 29 trials is approximately 1.2\%. While this outcome is unlikely it is not impossible.

    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation. Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6–1.5 gaps per simulation.

>>>>

    We applied the gap detection methods described in Appendix~\ref{sec:gap_detection}, along with the perturber identification methods from Appendix~\ref{sec:Perturber_Identification}. Results from all simulations are presented in Appendix~\ref{sec:gallery_of_gaps}. Once we identified the perturbers responsible for each gap, we conducted additional simulations to verify correct identification of the perturbers. These additional contain only the suspected clusters and were launched one-at-a-time to isolate the effects of a specific clusters. These results are shown in Fig.~\ref{fig:decomposition}. A summary of which perturber caused a gap and when is presented in Fig.~\ref{fig:histogram_impact_time}.

@@

    One striking result is that more than half of the observed gaps are caused by NGC 2808, which had a flyby about 200 Myr ago. In fact, 44 of the 50 simulations show a gap from this interaction. Upon examining their morphology (Appendix~\ref{sec:gallery_of_gaps}), we observe that these gaps are consistently similar in size and location within the leading tail across most simulations. This is no coincidence. As demonstrated by Sanders and Erkal, gap creation is a dispersion phenomenon, where stars affected by a perturber acquire slightly different orbital frequencies, causing them to gradually spread apart over time. Thus, smaller gaps are indicative of more recent encounters.

    It is also expected that a recent flyby would appear in most simulations, as uncertainties in a cluster's precise location increase as we look further back in time. In general, after enough time passes, different orbital solutions \textit{phase mix}, meaning a body could be found anywhere in its phase space given its Hamiltonian and initial conditions. The characteristic time scale is proportional to its orbital period. However, since the encounter between NGC~2808 and Palomar~5's stream occurred less than one orbital period ago for either cluster, the orbital solutions remain relatively similar, causing the flyby to appear consistently in most simulations.

    Next, we observe that most gaps occur in the leading tail. We believe this to be a coincidence, as we expect there to be no physical motivator to cause more gaps to occur in one tail than another. First, stars escape from the globular cluster's L1 and L2 Lagrange points at roughly equal rates, leading the same size and number of stars per tail. The two main differences each tail is that the particles find themselves at different locations on the orbital phase at a given point and time. The second is that the leading tail is about 400~pc closer to the Galactic center than the trailing tail. However, this difference is negligible compared to the orbital radius of 8-12kpc. As such, we expect the likelihood of a gap appearing in either tail to be equal, with the presence of one gap not influencing the probability of another.

    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation. However, 44 of these gaps are associated with NGC~2808. As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter. We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events. After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations. By integrating the binomial distribution, we compute that the probability of having at least 8 successes or fewer given 29 trials is approximately 1.2\%. While this outcome is unlikely it is not impossible. Additionally, by studying Fig.~\ref{fig:histogram_impact_time}, we notice that, although in much smaller numbers compared to NGC~2808, other pertubers cause gaps at the same time accross different simulations. These could also be violating the assumption of independent events.

    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation. Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6-1.5 gaps per simulation.
